---
title: "Spectrolab"
subtitle: "hyperspectral data wrangling with R"
author: "Jose Eduardo Meireles"
date: "jose.meireles@maine.edu"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library("spectrolab")
library("pls")

i_spec = read_spectra("../data/psr/", format = "sed")
i_spec = smooth(i_spec[ , 400:2400])
```

<style type="text/css">
.remark-slide-content {
    font-size: 140%;
}
</style>


#Slides and data are available on github:
#[https://github.com/meireles/learn_spectrolab](https://github.com/meireles/learn_spectrolab)

---
class: center, middle

```{r dev='pdf', fig.width = 14, fig.height = 11, echo=FALSE}
par(mar = c(6,6,4,4))
plot(i_spec, cex.lab = 2, cex.axis = 2, lwd = 3, col = c("blue", "orange"))
```

---

# Outline

1. Getting Started

2. Reading & Inspecting spectra

3. Visualizing spectra

4. Manipulating / Processing spectra

5. Exporting spectra

6. **Predicting leaf chemistry from spectra**

---
class: inverse, center, middle

# 1. Getting Started

---

# To start using spectrolab, you should:

### 1. Install it from either:

[CRAN](https://cran.r-project.org/web/packages/spectrolab/) (stable):

```{r eval=FALSE, tidy=FALSE}
install.packages("spectrolab")
```

[GitHub](https://github.com/meireles/spectrolab) (cutting edge):

```{r eval=FALSE, tidy=FALSE}
library("devtools")
devtools::install_github("meireles/spectrolab")
```

### 2. Load spectrolab into R:

```{r eval=TRUE, message=FALSE}
library("spectrolab")
```

---
class: inverse, center, middle
# 2. Reading & Inspecting spectra
---

# Read spectra from *instrument files*

```{r eval=TRUE, tidy=FALSE}
## Read the PSR spectra from a directory
psr_spec = read_spectra(path = "../data/psr/", format = "sed")
```

This results in an object of class **spectra**
```{r eval=TRUE, tidy=FALSE}
## Inspect spectra
psr_spec

## Find out the name of the two first individual spectra
names(psr_spec)[1:2]
```


---
# Read spectra from a *matrix*

```{r eval=TRUE, tidy=FALSE}
## Read CSV
csv = read.csv("../data/csv/spec_mat.csv", check.names = FALSE, as.is = TRUE)

## Look at the matrix
csv[1:3, 1:5]

## Convert the matrix to spectra
csv_spec = as.spectra(x = csv, name_idx = 1, meta_idxs = 2)

## and look at the spectra
csv_spec
```

---
class: inverse, center, middle
# 3. Visualizing spectra
---

# Plot spectra

### Statically with *plot*

```{r eval=TRUE, dev='pdf', fig.width = 10, fig.height = 4.5}
par(mfrow = c(1, 2))
plot(psr_spec, lwd = 2, col = c("orange", "darkgreen", "purple"))
plot(csv_spec, xlab = "Wavelength nm", main = "My Spec")
```

---
# Plot spectra

### Interactively with *plot_interactive*

```{r eval=FALSE}
# Demo
plot_interactive(psr_spec)
```

---
class: inverse, center, middle
# 4. Manipulating / Processing spectra
---

# Manipulating spectra: e.g. subset

```{r eval=TRUE, dev='pdf', fig.width = 5, fig.height = 5, error=TRUE}

## Subset spectra like if they were matrices
psr_spec_trim = psr_spec[1:2 , 400:700]
psr_spec_trim

## Except that you **MUST** subset bands by their wavelength, **not** by position!
psr_spec[ , 1:10]
```

---

# Processing spectra: e.g. smooth

```{r eval=TRUE, dev='pdf', fig.width = 9, fig.height = 4.5, error=FALSE, warning=FALSE, message=FALSE}
## Smooth spectra with a moving average
psr_spec_smooth = smooth(psr_spec,
                         method = "moving_average", n = 20)

## Plot them
par(mfrow = c(1, 2), mar = c(7, 5, 1, 1))
plot(psr_spec[1, 360:700])
plot(psr_spec_smooth[1, 360:700])
```

---
class: inverse, center, middle
# 5. Exporting spectra
---

# Convert spectra into other objects

Convert spectra into **matrix** or **data.frame** in order to export or use them in regressions, ordinations, etc.

```{r eval=TRUE, dev='pdf', fig.width = 5, fig.height = 5, error=TRUE}

## Matrix
my_data_as_matrix = as.matrix(psr_spec)

## Data Frame
my_data_as_df     = as.data.frame(psr_spec)

## Quick peek
my_data_as_df[1:3, 1:4]
```

---
class: inverse, center, middle
# 6. Predicting leaf chemistry from spectra
---

# Predicting leaf chlorophyll from spectra

```{r eval=TRUE, dev='pdf', fig.width = 6, fig.height = 6, error=TRUE}
## Load the pls library
library("pls")

## Convert the spectra with the chloro metadata into a data frame
dat  = as.data.frame(csv_spec)

## Remove useless columns and fix chloro class (factor to numeric)
dat["sample_name"] = NULL 
dat["chlorophyll"] = as.numeric(as.character(dat[ , "chlorophyll"]))

## Divide data into "training" and "validation"
training_data   = dat[1:50,   ]
validation_spec = dat[51:100, -1]
validation_chlo = dat[51:100, 1 ]
```
---

```{r eval=TRUE, dev='pdf', fig.width = 6, fig.height = 6}
## Fit the model
fit = plsr(chlorophyll ~ ., data = training_data, ncomp = 5)

## Predict chlorophyll from spectra!
prediction = predict(object = fit, newdata = validation_spec, ncomp = 5)

## Plot truth against predictions
plot(validation_chlo, prediction, cex = 1.2, pch = 16)
```
